---
title: "Writing Functions in R"
subtitle: "RStudio Instructor Certification Teaching Exam"
author: "Sharla Gelfand"
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: ["xaringan-themer.css", "pull.css"]
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: 16:9
---
class: center

```{r, include = FALSE}
library(flair)

library(xaringanthemer)
style_duo(
  primary_color = "#1F4257",
  secondary_color = "#F97B64",
  text_slide_number_font_size = "0em"
)
```

### You already know how to use functions!

--

.pull-left[
Reading a CSV...

`read_csv()`
]

--

.pull-right[
Calculating a mean...

`mean()`
]

--

.pull-left[
Making a plot...

`ggplot() + 
  geom_point()`
]

--

.pull-right[
Fitting a model...

`lm(y~x)`
]

--

Functions help you organize and reuse code.

Now let's write our own! 

???

As an R user, you already know how to use functions! Functions enable everything we do in R

Functions allow you to better organize and reuse your code - so today we're going to learn how to write our own.

---

I love to cook!

Americans give measurements in ounces.

--

```{r}
8 * 28.349
```

--

```{r}
16 * 28.349
``` 

--

If I only had an R function instead of doing the same thing over and over

---

```{r}
convert_ounces_to_grams <- function(ounces) {
  ounces * 28.349
}
```

```{r}
convert_ounces_to_grams(ounces = 8)
convert_ounces_to_grams(ounces = 16)
```

---

# Components of a function

```{r initial-function, include = FALSE}
convert_ounces_to_grams <- function(ounces) {
  ounces * 28.349
}
```

```{r, echo = FALSE}
decorate("initial-function") %>% 
  flair("convert_ounces_to_grams")
```

Name

---

# Components of a function

```{r, echo = FALSE}
decorate("initial-function") %>% 
  flair_rx("(?<=\\().+?(?=\\))")
```

Arguments

---

# Components of a function

```{r, echo = FALSE}
decorate("initial-function") %>% 
  flair("ounces * 28.349")
```

Body

---

# Naming functions

Clearly and meaningfully - write code that is easier to understand!

`convert_ounces_to_grams`

--

I like **snake_case**, some people like **camelCase** - use whatever feels right!

???

Functions should be named clearly and meaningfully, to help yourself write (and read!) code that is easier to understand - the names are for people as much as for computers!

convert_oz_to_grams()

I like to use snake_case, but some people like camelCase - use whatever feels right for you!

---

# Naming functions

Be consistent!

`str_replace()`

`str_detect()`

???

If you're creating many functions, it's helpful to name them consistently

The stringr package is a great example of this

You may not know what you're looking for yet, but you know how it starts!

---
class: inverse, center, middle

# Time for an exercise!

--

Imagine you're writing a function to convert kilograms to pounds.

Which name would you choose for your function?

`weight_converter()`

`convert_kg_to_lb()`

`kgtolb()`

`Kilogram_ToPounds()`

---

# Arguments

Data: what the function works on

```{r data-argument, include = FALSE}
convert_ounces_to_grams(ounces = 8)
```

```{r, echo = FALSE}
decorate_chunk("data-argument") %>%
  flair("ounces = 8")
```

--

My kitchen scale doesn't have decimals! Can we round?

---

# Arguments

Details: how the computation is run

```{r details-argument, include = FALSE}
convert_ounces_to_grams <- function(ounces, digits) {
  round(ounces * 28.349, digits = digits)
}
```

```{r, echo = FALSE}
decorate_chunk("details-argument") %>%
  flair("digits)")
```


```{r}
convert_ounces_to_grams(ounces = 8, digits = 0)
convert_ounces_to_grams(ounces = 8, digits = 1)
```

---
# Argument order

Data first, details after
---

Exercise on arguments and orders

---

# Default arguments

A **default** value that your function uses for an argument if nothing is supplied

--

```{r default-arguments, include = FALSE}
convert_ounces_to_grams <- function(ounces, digits = 0) {
  round(ounces * 28.349, digits = digits)
}
```

```{r, echo = FALSE}
decorate_chunk("default-arguments") %>%
  flair("digits = 0")
```

--

```{r}
convert_ounces_to_grams(ounces = 8)
```

--

Set the default to the value you want most often!

--

If you're using another well-known function, it's best to have the defaults match!

```{r}
round(1.9)
```


---

# Calling arguments

---

# Function body

Controls all calculations and returns the last thing computed

```{r}
convert_ounces_to_grams <- function(ounces, digits = 0) {
  round(ounces * 28.349, digits = digits)
}
```

We don't have to tell the function to return `round()`... it just does

HIGHLIGHT THE INSIDE

---

# Function body

Split long functions into multiple steps using **intermediate variables**.

These only exist while the code in the function is being run!

---
class: inverse, middle, center

# Live Coding Time!

---

# Function body

If you need to return something early???

---

# Wrapping up

```{r}
convert_ounces_to_grams <- function(ounces, digits = 0) {
  round(ounces * 28.349, digits = digits)
}
```

Functions are **named** meaningfully to help you write easier to understand code.

They take **arguments** which dictate what data your function is run on and the details of how it's run.

And they have a body, which can contain complicated calculations, but is self contained.
