---
title: "Writing Functions in R"
subtitle: "RStudio Instructor Certification Teaching Exam"
author: "Sharla Gelfand"
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: ["xaringan-themer.css", "pull.css"]
    nature:
      countIncrementalSlides: false
      ratio: 16:9
---
class: center

```{r, include = FALSE}
library(flair)

library(xaringanthemer)
style_mono_accent(
  base_color = "#1F4257",
  text_slide_number_font_size = "0em"
)
```

### You already know how to use functions!

--

.pull-left[
Reading a CSV...

`read_csv()`
]

--

.pull-right[
Calculating a mean...

`mean()`
]

--

.pull-left[
Making a plot...

`ggplot() + geom_point()`
]

--

.pull-right[
Fitting a model...

`lm(y~x)`
]

--

Functions help you organize and reuse code.

---

# Why functions?

1. Make your code easier to read, and focus on the big picture

2. Easier to make changes

3. Reduce mistakes from copy and pasting code

---

# Why functions?

.pull-left[
```{r}
mean(1:10)
mean(1:5)
```
]

--

.pull-right[
```{r}
(1 + 2 + 3 + 4 + 5 + 6 + 7 + 8 + 9 + 10) / 10
(1 + 2 + 4 + 5) / 5
```
]


---

# Writing a function

I love to cook!

Americans give measurements in ounces.

--

```{r}
8 * 28.349
```

--

```{r}
16 * 28.349
``` 

--

If I only had an R function instead of doing the same thing over and over

---

# Writing a function

```{r initial-function, include = FALSE}
convert_ounces_to_grams <- function(ounces) {
  ounces * 28.349
}
```

```{r, echo = FALSE}
decorate("initial-function")
```

--

```{r}
convert_ounces_to_grams(ounces = 8)
```

--

```{r}
convert_ounces_to_grams(ounces = 16)
```

---

# Function components

```{r, echo = FALSE}
decorate("initial-function") %>% 
  flair("convert_ounces_to_grams", background = "#ffff7f")
```

A function has a <span style = "background-color: #ffff7f">**name**</span>.

---

# Function components

```{r, echo = FALSE}
decorate("initial-function") %>% 
  flair("convert_ounces_to_grams", background = "#ffff7f") %>% 
  flair_rx("(?<=\\().+?(?=\\))", background = "#ffbf7f")
```

A function has a <span style = "background-color: #ffff7f">**name**</span>.

It takes <span style = "background-color: #ffbf7f">**arguments**</span>.

---

# Function components

```{r, echo = FALSE}
decorate("initial-function") %>% 
  flair("convert_ounces_to_grams", background = "#ffff7f") %>% 
  flair_rx("(?<=\\().+?(?=\\))", background = "#ffbf7f") %>%
  flair("ounces * 28.349", background = "#ff7f7f")
```

A function has a <span style = "background-color: #ffff7f">**name**</span>.

It takes <span style = "background-color: #ffbf7f">**arguments**</span>.

The <span style = "background-color: #ff7f7f">**body**</span> contains code for calculations - whatever you're trying to achieve!

---

# Naming functions

Clearly, meaningfully, unambiguously - write code that is easier to understand!

--

`convert_ounces_to_grams`

--

**snake_case**: words separated by underscores

**camelCase**: uppercase first letter of words

Use whatever feels right!

---
class: center, middle

### Time for an exercise!

---

# Arguments

Inputs that control *what* the function runs on

```{r data-argument-8, include = FALSE}
convert_ounces_to_grams(ounces = 8)
```

```{r, echo = FALSE}
decorate_chunk("data-argument-8") %>%
  flair("ounces = 8")
```

--

```{r data-argument-16, include = FALSE}
convert_ounces_to_grams(ounces = 16)
```

```{r, echo = FALSE}
decorate_chunk("data-argument-16") %>%
  flair("ounces = 16")
```

---

# Arguments

Additional arguments to control the *how*

```{r details-argument, include = FALSE}
convert_ounces_to_grams <- function(ounces, digits) {
  round(ounces * 28.349, digits = digits)
}
```

```{r, echo = FALSE}
decorate_chunk("details-argument") %>%
  flair("digits") %>%
  flair("digits = digits")
```

--

```{r}
convert_ounces_to_grams(ounces = 8, digits = 0)
```

```{r}
convert_ounces_to_grams(ounces = 8, digits = 1)
```

---

# Default arguments

Arguments can have a **default** - used if nothing is supplied.

No default? You have to give something!

### Live coding time!

--

```{r default-arguments, include = FALSE}
convert_ounces_to_grams <- function(ounces, digits = 0) {
  round(ounces * 28.349, digits = digits)
}
```

```{r, echo = FALSE}
decorate_chunk("default-arguments") %>%
  flair("digits = 0")
```

--

```{r}
convert_ounces_to_grams(ounces = 8)
```

```{r, error = TRUE}
convert_ounces_to_grams()
```

---

# Default arguments

Set the default to the value you use most.

--

Match defaults to *existing* functions

```{r}
round(1.9)
```

---

# Function body

The **body** of a function contains all calculations - hopefully you know what to put here!   

```{r body, include = FALSE}
convert_ounces_to_grams <- function(ounces, digits = 0) {
  round(ounces * 28.349, digits = digits)
}
```

```{r, echo = FALSE}
decorate_chunk("body") %>%
  flair("round(ounces * 28.349, digits = digits)")
```

---

# Function body

**Self-contained** - writing comments, variables, just like the rest of your code!

### Live coding time!

--

```{r body-complicated, include = FALSE}
convert_ounces_to_grams <- function(ounces, digits = 0) {
  # Convert to grams
  grams <- ounces * 28.349
  
  # Round grams
  round(grams, digits = digits)
}
```

```{r, echo = FALSE}
decorate_chunk("body-complicated") %>%
  flair("grams <- ounces * 28.349") %>%
  flair("# Apply rounding")
```

---

# What comes out of a function?

The last calculation is returned - don't have to tell your function what to do.

```{r body-return, include = FALSE}
convert_ounces_to_grams <- function(ounces, digits = 0) {
  # Convert to grams
  grams <- ounces * 28.349
  
  # Round grams
  round(grams, digits = digits)
}
```

```{r, echo = FALSE}
decorate_chunk("body-return") %>%
  flair("round(grams, digits = digits)")
```


---

# What comes out of a function?

Output something explicitly with `return()`

```{r body-return-explicit, include = FALSE}
convert_ounces_to_grams <- function(ounces, digits = 0) {
  # Convert to grams
  grams <- ounces * 28.349
  
  # Round grams
  return(round(grams, digits = digits))
}
```

```{r, echo = FALSE}
decorate_chunk("body-return-explicit") %>%
  flair("return")
```

---

# Returning early

You don't always need to run all the calculations.

--

What's 0 ounces?

--

```{r, return-early, include = FALSE}
convert_ounces_to_grams <- function(ounces, digits = 0) {
  # 0 ounces is 0 grams
  if (ounces == 0) {
    return(0)
  }
  
  # Convert to grams
  grams <- ounces * 28.349
  
  # Round grams
  round(grams, digits = digits)
}
```


```{r, echo = FALSE}
decorate_chunk("return-early") %>%
  flair("if (ounces == 0) {
    return(0)
  }")
```

---

# Functions

```{r}
convert_ounces_to_grams <- function(ounces, digits = 0) {
  # 0 ounces is 0 grams
  if (ounces == 0) {
    return(0)
  }
  
  # Convert to grams
  grams <- ounces * 28.349
  
  # Round grams
  round(grams, digits = digits)
}
```

Functions help you write code that is **clearer**, with less **repetition** and less **errors**.

They are **named** meaningfully so your code is easier to understand.

They take **arguments** which dictate what data your function is run on and the details of how it's run.

They have a **body**, which can contain complicated calculations, but are self-contained.
